---
title: "Safe Passage Brain Responses"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE, include = FALSE}
packs <-
  c("dplyr",
    "ggplot2",
    "lmerTest")
lapply(packs, require, character.only = TRUE)
```

```{r}
eeg_data <- read.csv("Data/csv/all_meas_new.csv") %>%
  mutate_if(is.character, function(x)
    gsub(
      pattern = "'",
      replacement = "",
      x = x
    ))
head(eeg_data)
```

# Clinical & Demographic variables

```{r}
all_vars <- read.csv("Data/csv/SafePassageSummaryScores.csv") %>%
  mutate(ELC = rowSums(.[c(139, 144, 149)])) # Early Learning Composite (ELC) = VisualReception_Raw, ReceptiveLanguage_RawScore, ExpressiveLanguage_RawScore
```

```{r}
melc <- mean(all_vars$ELC, na.rm = TRUE)
selc <- sd(all_vars$ELC, na.rm = TRUE)
vsm <-
  mean((
    all_vars$RecpetiveLanguage_Tscore + all_vars$ExpressiveLanguage_Tscore
  ) / 2,
  na.rm = TRUE
  )
vss <-
  sd((
    all_vars$RecpetiveLanguage_Tscore + all_vars$ExpressiveLanguage_Tscore
  ) / 2,
  na.rm = TRUE
  )
vrm <- mean(all_vars$VisualReception_Tscore, na.rm = TRUE)
vrs <- sd(all_vars$VisualReception_Tscore, na.rm = TRUE)

all_vars <- all_vars %>%
  mutate(
    ELC.st = 100 + (ELC - melc) * (15 / selc),
    FIQ_ELC = ifelse(is.na(FSIQ), ELC.st, FSIQ),
    
    VS = 100 + (((RecpetiveLanguage_Tscore + ExpressiveLanguage_Tscore) /
                   2
    ) - vsm) * (15 / vss),
    VR = 100 + (VisualReception_Tscore - vsm) * (15 / vrs),
    
    VCI_alt = ifelse(is.na(VCI), VS, VCI),
    PRI_alt = ifelse(is.na(PRI), VR, PRI),
    
    age.group = ifelse(Age_yrs < 6, 1, ifelse(
      Age_yrs < 11 & Age_yrs >= 6, 2, ifelse(Age_yrs >= 11, 3, NA)
    ))
  ) 
```

```{r}
cl_demo_vars <- all_vars[!duplicated(all_vars$Subject),] %>%
  select(
    Subject,
    Gender,
    Age_yrs,
    #age.group,
    FIQ_ELC,
    # COLIFE,
    VCI_alt,
    PRI_alt,
    # VisualReception_Tscore,
    # RecpetiveLanguage_Tscore,
    # ExpressiveLanguage_Tscore,
    #OLIFE_Mother,
    # maritalStatus,
    #Employed..0.no.1.yes.,
    #educationYears,
    #BMI...kg.m_.,
    household_income,
    # typeHouse,
    #phone..0.no.1.yes.,
    #water..0.no.1.yes.,
    #toilet..0.no.1.yes.,
    # Risk_TOTAL,
    # Protective_TOTAL,
    #Edinburgh.score,
    #PARITY..life.birth.,
    #CigsPerDayPreg,
    #TotalStdDrinksPreg,
    #Preterm...0.no.1.yes.,
    SDQ_tot,
    CARS_Total_Raw
  ) %>%
  rename(id = Subject) %>%
  mutate(id = gsub(
    pattern = "002-BISH-",
    replacement = "P",
    x = id
  ))
```

# Session info from tracker

```{r}
tracker <- read.csv("Data/csv/tracker.csv")
tracker <- tracker[!duplicated(tracker$id),] %>%
  select(-c(17:21)) %>%
  mutate_if(is.character, function(x)
    gsub(
      pattern = " ",
      replacement = "",
      x = x
    )) %>%
  select(id, age_ymd, time_capping_m)
```

# ET Data

```{r}
et_data <- read.csv("Data/csv/et_data.csv") %>%
  mutate(cond = ifelse(event == "FACE_EN.mp4", "social", "non-social")) %>%
  group_by(id, cond) %>%
  summarise_at(.vars = 6:11, mean)
```

# Merge all datasets

```{r}
eeg_data_m <- eeg_data %>%
  left_join(cl_demo_vars, by = "id") %>%
  left_join(et_data, by = c("id", "cond")) %>%
  left_join(tracker, by = "id")
ds <- filter(eeg_data_m, n.clean.ep > 20 & foof.rsq > 0.95)
```

# Descriptive Statistics

```{r}
summ.all <- ds %>%
  filter(!if_all(Gender:CARS_Total_Raw, is.na)) %>%
  group_by(id, Gender, Age_yrs, VCI_alt, PRI_alt, FIQ_ELC, CARS_Total_Raw, SDQ_tot, household_income, time_capping_m) %>%
  summarise(et_looking_prop = mean(et_looking_prop, na.rm=TRUE)) %>%
  as.data.frame() %>%
  summarise(N=n(),
            `N fem`=sum(Gender=="Female", na.rm = TRUE),
            `M:F`=round(sum(Gender=="Male", na.rm = TRUE)/sum(Gender=="Female", na.rm = TRUE), 2),
            `Age (SD)`= paste0(round(mean(Age_yrs, na.rm=TRUE), 2),
                                 " (",
                                 round(sd(Age_yrs, na.rm=TRUE), 2),
                                 ", ",
                                 round(max(Age_yrs, na.rm=TRUE)-min(Age_yrs, na.rm=TRUE), 2),
                                 ")"),
            `VCI (SD)`= paste0(round(mean(VCI_alt, na.rm=TRUE), 2),
                                 " (",
                                 round(sd(VCI_alt, na.rm=TRUE), 2),
                                 ", ",
                                 round(max(VCI_alt, na.rm=TRUE)-min(VCI_alt, na.rm=TRUE), 2),
                                 ")"),
            `PRI (SD)`= paste0(round(mean(PRI_alt, na.rm=TRUE), 2),
                                 " (",
                                 round(sd(PRI_alt, na.rm=TRUE), 2),
                                 ", ",
                                 round(max(PRI_alt, na.rm=TRUE)-min(PRI_alt, na.rm=TRUE), 2),
                                 ")"),
            ELC =  paste0(round(mean(FIQ_ELC, na.rm=TRUE), 2),
                                 " (",
                                 round(sd(FIQ_ELC, na.rm=TRUE), 2),
                                 ", ",
                                 round(max(FIQ_ELC, na.rm=TRUE)-min(FIQ_ELC, na.rm=TRUE), 2),
                                 ")"),
            `CARS (SD)`= paste0(round(mean(CARS_Total_Raw, na.rm=TRUE), 2),
                                 " (",
                                 round(sd(CARS_Total_Raw, na.rm=TRUE), 2),
                                 ", ",
                                 round(max(CARS_Total_Raw, na.rm=TRUE)-min(CARS_Total_Raw, na.rm=TRUE), 2),
                                 ")"),
            `SDQ (SD)`= paste0(round(mean(SDQ_tot, na.rm=TRUE), 2),
                                 " (",
                                 round(sd(SDQ_tot, na.rm=TRUE), 2),
                                 ", ",
                                 round(max(SDQ_tot, na.rm=TRUE)-min(SDQ_tot, na.rm=TRUE), 2),
                                 ")"),
            household_income = paste0(round(mean(household_income, na.rm=TRUE), 2),
                               " (",
                               round(sd(household_income, na.rm=TRUE), 2),
                               ", ",
                               round(max(household_income, na.rm=TRUE)-min(household_income, na.rm=TRUE), 2),
                               ")"),
            et_looking_prop = paste0(round(mean(et_looking_prop, na.rm=TRUE), 2),
                               " (",
                               round(sd(et_looking_prop, na.rm=TRUE), 2),
                               ", ",
                               round(max(et_looking_prop, na.rm=TRUE)-min(et_looking_prop, na.rm=TRUE), 2),
                               ")"),
            time_capping_m = paste0(round(mean(time_capping_m, na.rm=TRUE), 2),
                               " (",
                               round(sd(time_capping_m, na.rm=TRUE), 2),
                               ", ",
                               round(max(time_capping_m, na.rm=TRUE)-min(time_capping_m, na.rm=TRUE), 2),
                               ")")
            ) 
tab_sum <- summ.all %>% t() %>% as.data.frame() %>%
  tibble::rownames_to_column(" ")
tab_sum
```

```{r}
ds %>%
  filter(!if_all(Gender:CARS_Total_Raw, is.na)) %>%
  distinct(id, Age_yrs, FIQ_ELC) %>%
  filter(Age_yrs>6) %>%
  summarise(N=n(),
            paste0(round(mean(FIQ_ELC, na.rm=TRUE), 2),
                                 " (",
                                 round(sd(FIQ_ELC, na.rm=TRUE), 2),
                                 ", ",
                                 round(max(FIQ_ELC, na.rm=TRUE)-min(FIQ_ELC, na.rm=TRUE), 2),
                                 ")"))
ds %>%
  filter(!if_all(Gender:CARS_Total_Raw, is.na)) %>%
  distinct(id, Age_yrs, FIQ_ELC) %>%
  filter(Age_yrs<=6) %>%
  summarise(N=n(),
            paste0(round(mean(FIQ_ELC, na.rm=TRUE), 2),
                                 " (",
                                 round(sd(FIQ_ELC, na.rm=TRUE), 2),
                                 ", ",
                                 round(max(FIQ_ELC, na.rm=TRUE)-min(FIQ_ELC, na.rm=TRUE), 2),
                                 ")"))
ds %>%
  filter(!if_all(Gender:CARS_Total_Raw, is.na)) %>%
  group_by(cond) %>%
  summarise(paste0(round(mean(et_looking_prop, na.rm=TRUE), 2),
                                 " (",
                                 round(sd(et_looking_prop, na.rm=TRUE), 2),
                                 ", ",
                                 round(max(et_looking_prop, na.rm=TRUE)-min(et_looking_prop, na.rm=TRUE), 2),
                                 ")"))
```

Filtering:

-   Number of clean epochs: We use 20 trials as a minimum amount, so
    exclude data with less than 20 trials.

-   r-squared for the fit of the model from Fitting of one over f
    (FOOOF): a value below .95 is considered bad, so exclude for
    analyses with FOOOF measures: FOOOFIntercept, FOOOFSlope,
    PeriodicPeakFrequency, PeriodicPeakAmplitude, PeriodicPower (This
    value is identical across frequency bands of interest because it is
    fitted from 1 to 30Hz).

# N kids showing peaks in alpha/range as bar graph

Purpose: information on inclusion

```{r}
eeg_data_m %>%
  filter(!is.nan(per.pk.fr)) %>%
  select(id, f.b, reg, cond) %>%
  group_by(f.b, reg, cond) %>%
  ggplot(aes(f.b)) +
  geom_bar(aes(fill = reg), width = 0.5) +
  facet_grid( ~ cond) +
  theme(axis.text.x = element_text(angle = 65, vjust = 0.6)) +
  labs(
    title = "",
    subtitle = "",
    x = "Frequency Band",
    y = "N of Trials",
    fill = "Region"
  ) +
  theme_bw()

```

# Data Quantity

DVs:

1.  N of clean epochs

2.  Fit error

Predictors:

-   Condition

-   Environmental variables:

    -   visual attention

    -   capping time

-   Phenotype:

    -   Age

    -   Gender

    -   FIQ

    -   SDQ

    -   CARS

    -   Household income

    -   Condition \* Age

## DV = N of clean epochs

This measure is identical across frequency bands and regions (alpha and
Frontal have been selected, but theta will retain same result).

```{r}
m_nce <-
  lmer(
    data = filter(ds, f.b == "theta" & reg == "Frontal"),
    n.clean.ep ~ cond * Age_yrs + et_looking_prop + time_capping_m +
      Gender + FIQ_ELC + SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 | id)
  )
summ_nce <- m_nce %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x)
    round(x, 4)) %>%
  flextable::flextable()
summ_nce
```
The number of clean epochs is significantly influenced by the Eye-Tracking Looking Proportion (positive).

## DV = Foof fit error

This measure is identical across frequency bands and regions (alpha and
Frontal have been selected, but theta will retain same result).

```{r}
m_rsq <-
  lmer(
    data = filter(eeg_data_m, f.b == "alpha" & reg == "Frontal"),
    foof.rsq ~ cond * Age_yrs + et_looking_prop + time_capping_m +
      Gender + FIQ_ELC + SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 | id)
  )
summ_rsq <- m_rsq %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x)
    round(x, 4)) %>%
  flextable::flextable()
summ_rsq
```

The foof fit error is significantly influenced by Age (negative).

# Condition Estimates

DVs:

1.  Canonical power theta

2.  Canonical power alpha

3.  1/f intercept

4.  1/f slope

5.  Frequency of periodic peak in theta range

6.  Amplitude of periodic peak in theta range

7.  Frequency of periodic peak in alpha range

8.  Frequency of periodic peak in alpha range

9.  Periodic power theta

10. Periodic power alpha

Predictors:

-   Number of clean epochs

-   Age

-   Gender

-   FIQ

-   SDQ

-   CARS

-   Household income

-   Condition \* Age

Each table includes the type of effect (fixed, random), grouping (just for random effects), term (meaning the predictor), estimate (meaning the coefficient, a measure of the unstandardized effect size for that
predictor), standard error, statistic (meaning the T-value or distance form 0) and p-value. A brief paragraph summarizes which predictors are significantly associated with the DV, and the direction of the association in parenthesis (positive/negative, e.g., a positive association means that for higher values of the predictor, the DV also takes higher values, while a negative association means that for higher
values of the predictor, the DV takes lower values).

## DV = Canonical Power Theta

```{r}
m_powT <-
  lmer(
    data = filter(eeg_data_m, n.clean.ep > 20 & f.b == "theta"),
    can.pow ~ n.clean.ep + cond * Age_yrs + Gender + FIQ_ELC +
      SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 + reg | id)
  )
summary(m_powT)
summ_powT <- m_powT %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x)
    round(x, 4)) %>%
  flextable::flextable()
summ_powT
```

The Canonical Theta Power is significantly influenced by the N of Clen Epochs (negative), Condition (in the social condition: positive), Age (negative), and the interaction of condition by age (negative).

## DV = Canonical Power Alpha

```{r}
m_powA <-
  lmer(
    data = filter(eeg_data_m, n.clean.ep > 20 & f.b == "alpha"),
    can.pow ~ n.clean.ep + cond * Age_yrs + Gender + FIQ_ELC +
      SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 + reg | id)
  )
summ_powA <- m_powA %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x)
    round(x, 4)) %>%
  flextable::flextable()
summ_powA
```

The Canonical Alpha Power is significantly influenced by Age (negative).

## DV = 1/f intercept

This measure is identical across frequency bands (alpha has been
selected, but theta will retain same result).

```{r}
m_1fint <-
  lmer(
    data = filter(eeg_data_m, n.clean.ep > 20 & foof.rsq > 0.95 &
                    f.b == "theta"),
    foof.int ~ n.clean.ep + cond * Age_yrs + Gender + FIQ_ELC +
      SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 + reg | id)
  )
summ_1fint <- m_1fint %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x)
    round(x, 4)) %>%
  flextable::flextable()
summ_1fint
```

The 1/f Intercept is significantly influenced by the N of Clean Epochs (negative) and Age (negative).

## DV = 1/f slope

This measure is identical across frequency bands (alpha has been
selected, but theta will retain same result).

```{r}
m_1fsl <-
  lmer(
    data = filter(eeg_data_m, n.clean.ep > 20 & foof.rsq > 0.95 &
                    f.b == "alpha"),
    foof.sl ~ n.clean.ep + cond * Age_yrs + Gender + FIQ_ELC +
      SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 + reg | id)
  )
summ_1fsl <- m_1fsl %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x)
    round(x, 4)) %>%
  flextable::flextable()
summ_1fsl
```

The 1/f Slope is significantly influenced by the N of Clean Epochs (negative).

## DV = Frequency of periodic peak in theta range

```{r}
m_ppfT <-
  lmer(
    data = filter(eeg_data_m, n.clean.ep > 20 & foof.rsq > 0.95 &
                    f.b == "theta"),
    per.pk.fr ~ n.clean.ep + cond * Age_yrs + Gender + FIQ_ELC +
      SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 + reg | id)
  )
summ_ppfT <- m_ppfT %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x)
    round(x, 4)) %>%
  flextable::flextable()
summ_ppfT
```

The frequency of periodic peak in theta range is significantly influenced by condition (in social condition: positive), FIQ (negative), CARS (negative), and the interaction of condition by age (negative).

## DV = Amplitude of periodic peak in theta range

```{r}
m_ppaT <-
  lmer(
    data = filter(eeg_data_m, n.clean.ep > 20 & foof.rsq > 0.95 &
                    f.b == "theta"),
    per.pk.amp ~ n.clean.ep + cond * Age_yrs + Gender + FIQ_ELC +
      SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 + reg | id)
  )
summ_ppaT <- m_ppaT %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x)
    round(x, 4)) %>%
  flextable::flextable()
summ_ppaT
```

The amplitude of periodic peak in theta range is significantly influenced by condition (in social condition: positive) and the interaction of condition by age (negative).

## DV = Frequency of periodic peak in alpha range

```{r}
m_ppfA <-
  lmer(
    data = filter(eeg_data_m, n.clean.ep > 20 & foof.rsq > 0.95 &
                    f.b == "alpha"),
    per.pk.fr ~ n.clean.ep + cond * Age_yrs + Gender + FIQ_ELC +
      SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 + reg | id)
  )
summ_ppfA <- m_ppfA %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x)
    round(x, 4)) %>%
  flextable::flextable()
summ_ppfA
```

The frequency of periodic peak in alpha range is significantly influenced by condition (positive), Age (positive) and the interaction of condition by age (negative).

## DV = Frequency of periodic peak in alpha range

```{r}
m_ppaA <-
  lmer(
    data = filter(eeg_data_m, n.clean.ep > 20 & foof.rsq > 0.95 &
                    f.b == "alpha"),
    per.pk.amp ~ n.clean.ep + cond * Age_yrs + Gender + FIQ_ELC +
      SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 + reg | id)
  )
summ_ppaA <- m_ppaA %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x)
    round(x, 4)) %>%
  flextable::flextable()
summ_ppaA
```

The amplitude of periodic peak in alpha range is significantly influenced by Age (negative) and household income (negative).

## DV = Periodic power theta

```{r}
m_ppT <-
  lmer(
    data = filter(eeg_data_m, n.clean.ep > 20 & foof.rsq > 0.95 &
                    f.b == "theta"),
    per.pow ~ n.clean.ep + cond * Age_yrs + Gender + FIQ_ELC +
      SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 + reg | id)
  )
summ_ppT <- m_ppT %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x)
    round(x, 4)) %>%
  flextable::flextable()
summ_ppT
```

The periodic power in theta range is significantly influenced by condition (in social: positive), age (negative) and the interaction of condition by age (negative).

## DV = Periodic power alpha

```{r}
m_ppA <-
  lmer(data = filter(eeg_data_m, n.clean.ep > 20 & foof.rsq > 0.95 & f.b == "alpha"),
    per.pow ~ n.clean.ep + cond * Age_yrs + Gender + FIQ_ELC +
      SDQ_tot + CARS_Total_Raw + scale(household_income) +
      (1 + reg | id)
  )
summ_ppA <- m_ppA %>%
  broom.mixed::tidy() %>%
  mutate_if(is.numeric, function(x) round(x, 4)) %>%
  flextable::flextable()
summ_ppA
```

The periodic power in alpha range is significantly influenced by the N of clean epochs (positive) and household income (negative).
